---
output:
  pdf_document:
    extra_dependencies: ["bm","url"]
---

STAT430 Homework #2: Due Friday, February 11, 2022. 
========================================================

#### Name: KEY

*********************************************************************************

For this homework, you *must* submit a pdf file to Canvas.  You can still submit handwritten solutions if you wish to solve some problems by hand.  However, I instead encourage you to learn LaTeX and write up your solutions in the R Markdown file.  I have not had any luck yet getting you paid acounts for RStudio cloud, but for now you can use the free version (\url{https://rstudio.cloud/}).  Let me know if you have problems with this.

# Question 1
Let $X_i$, $i=1,2,3$, be independent with N$(i,i^2)$ distributions.  For each of the following situations, use the $X_i$s to construct a statistic with the indicated distribution.  

1. chi-squared with 3 degrees of freedom

\textcolor{red}{
For $X \sim N(\mu, \sigma^2)$, $Z = \frac{X - \mu}{\sigma}$ is a standard normal random variable, and for $Z_i, \dots, Z_n \overset{\text{iid}}{\sim} N(0,1)$, $Z_i^2 + \dots + Z_n^2$ is a $\chi^2$ random variable with $n$ degrees of freedom. We can therefore construct a $\chi^2_3$ random variable with 
$$\left( \frac{X_1 - 1}{1} \right)^2 + \left( \frac{X_2 - 2}{2} \right)^2 + \left( \frac{X_3 - 3}{3} \right)^2$$
}

2. $t$ distribution with 2 degrees of freedom

\textcolor{red}{
A $t$ distributed random variable $T$ with $n$ degrees of freedom can be constructed with a standard normal random variable $Z$ and an independent $\chi^2_n$ random variable $W$ as
$$T = \frac{Z}{\sqrt{W/n}} \sim t_n$$
As before, we can construct a standard normal random variable using $X_1$ and an independent $\chi^2_2$ random variable with $X_2$ and $X_3$. Then we can get a $t_2$ distributed random variable with
$$\frac{\frac{X_1 - 1}{1}}{\sqrt{ \left[ \left( \frac{X_2 - 2}{2}\right)^2 + \left( \frac{X_3 - 3}{3} \right)^2 \right]\Big/2 }}$$
}

3. F distribution with 1 and 2 degrees of freedom

\textcolor{red}{
An $F_{n,m}$ distributed random variable $V$ can be constructed as the ratio of a $\chi^2_n$ divided by its degrees of freedom random variable $W$ and an independent $\chi^2_m$ random variable $U$ divided by its degrees of freedom:
$$V = \frac{W/n}{U/m} \sim F_{n,m}.$$
So,
$$\frac{\left( \frac{X_1 - 1}{1} \right)^2\Big/1}{\left( \left(\frac{X_2 - 2}{2}\right)^2 + \left(\frac{X_3 - 3}{3}\right)^2 \right)\Big/2} \sim F_{1,2}$$
}

# Question 2
Actuary students from CSU graduate and find a job paying an amount of money that is distributed $N(55000,2400^2)$, while CU students earning are distributed $N(55500,1200^2)$.

We can plot these distributions by creating a long vector of `x` values, here ranging from 50K to 60K, and evaluating each of the normal probability density functions using the `dnorm` function at these values of `x`. We then ask `R` to "connect the dots" to make a curve for each density.  Note that the functions for the normal distribution in `R` use take a mean and a standard deviation as arguments.

```{r}
x <- seq(50000,60000,length=1000) # vector of values from 50K - 60K
plot(x,dnorm(x,55000,2400),type="l",ylim=c(0,.00035)) # CSU students
lines(x,dnorm(x,55500,1200),type="l",col="blue") # CU students
```

We can also use `R` to calculate the probability that a normal random variable is within a given range using the `pnorm` function.  Let's look at the help file for this function.

```{r, eval = FALSE}
help(pnorm)
```

The default argument is `lower.tail=TRUE` indicating you will get the probability the normal random variable is less than the specified value `q`.  Thus, by default is it the cumulative distribution function. Use the `pnorm` command to answer the questions below.

1. You are offered an entry level job paying $53,800.  At which institution do you rank higher?  In other words, what percentile do you represent at each institution and which is higher?

```{r}
pnorm(53800, 55000, 2400) # Percentile for CSU Students
pnorm(53800, 55000, 1200) # Percentile for CU Students
```

\textcolor{red}{
\$53,800 is higher than 31\% of the salaries of CSU actuary graduates and 16\% of CU actuary graduates. Therefore, you rank higher among CSU actuary graduates.
}

2. Your friend is offered a job paying $58,000.  At which institution does your friend rank higher?  

```{r}
pnorm(58000, 55000, 2400) # Percentile for CSU Students
pnorm(58000, 55000, 1200) # Percentile for CU Students
```

\textcolor{red}{
\$58,000 is higher than 89\% of the salaries of CSU actuary graduates and 99\% of CU actuary graduates. Therefore, your friend ranks higher among CU actuary graduates.
}

# Question 3

 Suppose $X \sim \chi^2_n$ distribution.  Prove that $$\frac{X-n}{\sqrt{2n}} 
\rightarrow_d N(0,1) \; \; \text{ as $n \rightarrow \infty$}.$$

\textcolor{red}{
The central limit says that for an iid sample $X_1, \dots, X_n$ of random variables
$$\frac{\bar{X} - \mu}{\sigma/\sqrt{n}} \overset{d}{\longrightarrow} N(0,1).$$
$X$ can be expressed as the sum of $X_1, \dots, X_n \overset{\text{iid}}{\sim} \chi^2_1$ random variables: $X = \sum_{i=1}^n X_i$. Define $\bar{X} = \frac{1}{n}\sum_{i=1}^n X_i$. Using the knowledge that $E[X_i] = 1$ and Var$(X_i) = 2$, we know that
$$\frac{\bar{X} - 1}{2/\sqrt{n}} \overset{d}{\longrightarrow} N(0,1)$$
multiplying the numerator and denominator by $n$, we get
$$\frac{\sum_{i=1}^n X_i - n}{2\sqrt{n}}  = \frac{X - n}{2\sqrt{n}} \overset{d}{\longrightarrow} N(0,1).$$
}

# Question 4

Let $\bar{X}$ be the average of 16 iid standard normal random variables.  Find $c$ such that 
\[
  P(|\bar{X}| < c) = 0.5.
\]

\textcolor{red}{
We can find $c$ as follows:
$$P(|\bar{X}| < c) = P(-c < \bar{X} < c)$$
$$0.5 = P\left( \frac{-c - 0}{1/4} < \frac{\bar{X} - 0 }{1/4} < \frac{c - 0 }{1/4}  \right) = P\left( -4c < \frac{\bar{X} - 0 }{1/4} < 4c  \right).$$
$\frac{\bar{X} - 0 }{1/4}$ is $t_{15}$ distributed, which is symmetric so we can use R to find the lower 25th percentile and use this to find $c$: 
}

```{r}
qt(.25, 15)
```

\textcolor{red}{
So $-4c \approx -0.6912$ and $c \approx 0.1728$. 
}

# Question 5

If $X \sim F_{n,m}$, find the distribution of $X^{-1}$.  Explain your reasoning.

\textcolor{red}{
If $X \sim F_{n,m}$ we can express $X$ as 
$$X = \frac{W/n}{V/m}$$
where $W \sim \chi^2_n$ and $V \sim \chi^2_m$. Taking the reciprocal, we have
$$X^{-1} = \frac{V/m}{W/n}$$
which is again an $F$ distributed random variable, but with $m$ and $n$ degrees of freedom.
}

# Question 6

If $X \sim T_{n}$, find the distribution of $X^2$.  Explain your reasoning.

\textcolor{red}{
If $X \sim t_n$, then $X$ can be expressed as
$$X = \frac{Z}{\sqrt{W/n}}$$
where $Z \sim N(0,1)$ and $W \sim \chi^2_n$. Squaring $X$, we get
$$X^2 = \frac{Z^2}{W/n} = \frac{Z^2/1}{W/n}.$$
which is the ratio of a $\chi^2_1$ random variable (divided by its degrees of freedom) and a $\chi^2_n$ random variable (divided by its degrees of freedom). Therefore $X^2 \sim F_{1,n}$.
}